\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Research Proposal: Enhancing AIOps with Reinforcement Learning through OpenR Framework Integration}

\author{\IEEEauthorblockN{1\textsuperscript{st} Wentao Ye}
    % \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
    %     \textit{name of organization (of Aff.)}\\
    %     City, Country \\
    %     email address or ORCID}
}

\date{\today}

\maketitle

\begin{abstract}
    In IT Operations (ITOps), managing complex cloud environments requires intelligent systems that emulate human reasoning to detect, diagnose, and resolve issues efficiently. 
    AIOpsLab introduces a structured framework that divides IT operations into four main stages: Incident Detection, Triage after Localization, Root Cause Diagnosis, and Mitigation Resolution. 
    Each stage requires the agent to interact with the cloud environment iteratively to achieve its objectives. 

    The OpenR framework introduces a novel reinforcement learning (RL) suite designed to model reasoning-intensive tasks. 
    It structures reasoning processes as a sequence of steps (Q $\rightarrow$ \{R\} $\rightarrow$ A) modeled as a Markov Decision Process (MDP), offering a structured mechanism for reasoning that aligns with the demands of IT operations. 
    
    This proposal explores the integration of OpenR into the AIOpsLab to create an RL-driven solution for IT operations, leveraging OpenR's reasoning-focused approach to enhance performance across all operational stages.
\end{abstract}

\section{Background and Motivation}

\subsection{IT Operations Framework (AIOpsLab)}
AIOpsLab outlines IT operations as a structured process consisting of four stages:
\begin{enumerate}
    \item \textbf{Incident Detection}: Identifying anomalies or issues in the system.
    \item \textbf{Triage after Localization}: Narrowing down potential causes and prioritizing them.
    \item \textbf{Root Cause Diagnosis}: Pinpointing the exact cause of the issue.
    \item \textbf{Mitigation Resolution}: Implementing corrective actions to resolve the issue.
\end{enumerate}

Each stage involves iterative reasoning and interaction with the cloud environment, resembling human problem-solving. An RL agent capable of emulating this reasoning can significantly enhance the speed, accuracy, and efficiency of IT operations.

\subsection{The OpenR Framework}
The OpenR framework proposes a structured methodology for reasoning tasks, particularly in scenarios that require iterative problem-solving and question-answering. It defines the reasoning process as a sequence (Q $\rightarrow$ \{R\} $\rightarrow$ A), where:
\begin{itemize}
    \item \textbf{Q} represents the question or prompt initiating the reasoning process.
    \item \textbf{R} represents the intermediate reasoning steps taken to build toward a solution.
    \item \textbf{A} represents the final answer or solution derived from the reasoning steps.
\end{itemize}

By framing reasoning as a Markov Decision Process (MDP), OpenR provides flexibility in modeling both sequential and branching reasoning paths:
\begin{itemize}
    \item \textbf{States} represent the sequence of reasoning steps taken so far.
    \item \textbf{Actions} involve selecting the next reasoning step or final answer.
    \item \textbf{Policy} refers to the LLM's function for generating reasoning steps and answers.
    \item \textbf{Process Reward Model (PRM)} provides feedback on the quality of reasoning steps and final answers, guiding the LLM policy to optimize its output.
\end{itemize}

This combination of sequential and branching reasoning allows OpenR to explore diverse solutions while maintaining logical consistency, making it a robust framework for reasoning-intensive tasks.

\subsection{Motivation for Integration}
Both OpenR and AIOpsLab emphasize the importance of reasoning in task-solving. By integrating OpenR into AIOpsLab, we aim to:
\begin{enumerate}
    \item \textbf{Enhance Agent Performance}: Leverage OpenR's structured reasoning approach to improve performance in each operational stage.
    \item \textbf{Enable Adaptive Learning}: Use OpenR's RL methodology to continuously refine the agent's decision-making process during cloud interactions.
    \item \textbf{Create a Scalable Framework}: Develop a scalable methodology for training RL agents in dynamic cloud environments.
\end{enumerate}

\section{Objectives}
\begin{enumerate}
    \item \textbf{Framework Integration}:
    \begin{itemize}
        \item Adapt OpenR's RL suite for IT operations tasks within the AIOpsLab framework.
        \item Leverage the Q $\rightarrow$ \{R\} $\rightarrow$ A structure to align reasoning processes with IT operations workflows.
    \end{itemize}
    
    \item \textbf{Performance Optimization}:
    \begin{itemize}
        \item Investigate the impact of OpenR's reasoning-based RL approach on agent performance across the four IT operations stages.
    \end{itemize}
    
    \item \textbf{Online Learning}:
    \begin{itemize}
        \item Implement online policy learning during inference to enable real-time performance improvements.
    \end{itemize}
    
    \item \textbf{Scalability and Robustness}:
    \begin{itemize}
        \item Develop scalable training pipelines to generalize the RL framework for diverse IT operations scenarios.
    \end{itemize}
\end{enumerate}

\section{Research Methodology}

\subsection{Framework Design}
\begin{enumerate}
    \item \textbf{Agent Architecture}:
    \begin{itemize}
        \item Extend the AIOpsLab agent to incorporate OpenR's RL suite.
        \item Structure agent reasoning using the Q $\rightarrow$ \{R\} $\rightarrow$ A sequence for step-by-step problem-solving.
    \end{itemize}
    
    \item \textbf{Process Reward Model (PRM)}:
    \begin{itemize}
        \item Adapt OpenR's PRM to evaluate intermediate steps and final outputs for IT operations tasks.
        \item Explore fine-tuning the PRM using IT operations datasets to provide domain-specific feedback.
    \end{itemize}
    
    \item \textbf{Interactive Environment}:
    \begin{itemize}
        \item Develop a simulated cloud environment where the agent can practice and refine its reasoning processes.
    \end{itemize}
\end{enumerate}

\subsection{Online Policy Learning}
\begin{itemize}
    \item Enable online policy learning during inference to dynamically refine the agent's behavior as it interacts with the cloud environment.
    \item Use retrieval-augmented generation (RAG) to provide the agent with context-specific knowledge.
\end{itemize}

\subsection{Dataset Preparation}
\begin{itemize}
    \item Utilize logs, incidents, and historical data from IT operations to create a reasoning-oriented dataset.
    \item Incorporate \textbf{human-in-the-loop annotation} to ensure data quality and relevance.
\end{itemize}

\subsection{Evaluation Metrics}
In addition to the metrics outlined in the AIOpsLab paper, we want to propose a \textbf{custom metric} to evaluate the performance of the integrated OpenR-AIOps framework.

\section{Key Challenges}

\subsection{Dataset and Domain Adaptation}
Adapting OpenR's reasoning framework to IT operations tasks presents two interconnected challenges:
\begin{enumerate}
    \item \textbf{Dataset Preparation}: Unlike OpenR's math-reasoning focus, IT operations lack a well-structured dataset for process-level supervision. Creating a dataset that accurately reflects the reasoning processes involved in IT operations—such as diagnosing root causes or resolving incidents—is costly and complex.
    \item \textbf{Framework Modification}: The OpenR framework relies on Process Reward Models (PRMs) fine-tuned on task-specific datasets (e.g., OmegaPRM for math reasoning). For IT operations, the lack of pre-existing high-quality datasets for process supervision means we may need to modify the PRM design or even the reasoning structure itself (Q $\rightarrow$ \{R\} $\rightarrow$ A) to better suit IT-specific challenges.
\end{enumerate}

\subsection{Computational Costs}
Scaling RL training for reasoning-intensive tasks like IT operations is resource-intensive. OpenR's RL suite requires significant computational resources for training its models, especially when incorporating branching reasoning paths or running simulations in dynamic cloud environments.

\section{Expected Contributions}
\begin{enumerate}
    \item A novel integration of the OpenR framework into the AIOpsLab for IT operations tasks.
    \item An RL-driven methodology for enhancing reasoning and decision-making in dynamic cloud environments.
    \item A benchmark for evaluating RL-based solutions in IT operations, emphasizing reasoning quality and task efficiency.
\end{enumerate}

\section{Conclusion}
Integrating the OpenR framework into AIOpsLab presents a transformative opportunity to enhance IT operations by leveraging reasoning-focused reinforcement learning. This research will contribute to developing more intelligent, adaptive, and efficient autonomous systems for managing complex cloud environments.

\begin{thebibliography}{00}
    \bibitem{b1} Y. Chen et al., “AIOpsLab: A Holistic Framework to Evaluate AI Agents for Enabling Autonomous Clouds”.
    \bibitem{b2} M. Shetty et al., “Building AI Agents for Autonomous Clouds: Challenges and Design Principles,” Jul. 31, 2024, arXiv: arXiv:2407.12165. doi: 10.48550/arXiv.2407.12165.
    \bibitem{b3} L. Zheng et al., “Judging LLM-as-a-Judge with MT-Bench and Chatbot Arena,” Dec. 24, 2023, arXiv: arXiv:2306.05685. doi: 10.48550/arXiv.2306.05685.
    \bibitem{b4} J. Wang et al., “OpenR: An Open Source Framework for Advanced Reasoning with Large Language Models,” Oct. 12, 2024, arXiv: arXiv:2410.09671. doi: 10.48550/arXiv.2410.09671.
\end{thebibliography}

\end{document}
